---
title: "JAGS Assignment"
author: "Yiru Fei"
date: "4/5/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjags)
library(runjags)
library(tidyverse)
```

## Exercise 8.1

```{r cars}
data(mtcars)
y = select(mtcars,vs)
s = c( rep("Apple",6) , rep("Blurberry",12) , rep("Cherry",14))
data = data.frame(y=y,s=s)
colnames(data)[1] = c("y")
rownames(data) = c(1:32)
source("DBDA2E-utilities.R")
source("Jags-Ydich-XnomSsubj-MbernBeta.R")
```
```{r,warning=FALSE}
mcmc = genMCMC( data=data , numSavedSteps=10000)
parameterNames = varnames(mcmc)
for ( parName in parameterNames ) {
diagMCMC( codaObject=mcmc , parName=parName )
}
smryMCMC( mcmc , compVal=NULL , compValDiff=0.0 )
plotMCMC( mcmc, data=data , compVal=NULL , compValDiff=0.0 )
```
Looking at the diagonal panels, we see that the HDI for theta[1] is wider than the HDI for theta[2], and the MCSE for theta[2] is wider than the MCSE for theta[3], because the data for theta[1] are far fewer than the data for theta[2]. 

## Exercise 8.2
```{r}
smryMCMC( mcmc , compVal=0.5 , rope=c(0.45,0.55) ,compValDiff=0.0 , ropeDiff = c(-0.05,0.05) )
```  
Each row represents the parameter or parameter difference. The columns labelled Mean, Median, and Mode are the corresponding values of parameter. The ESS is the effective sample size. The next three columns indicate the probability mass of the HDI, the lower limit of the HDI, and the upper limit. The comparison value is specified in the argument as 0.5. The next column indicates the percentage of the posterior that is greater than the comparison value (PcntGtCompVal). The ROPE columns repeat the specifications in the arguments. The final three columns indicate the percentage of the posterior distribution that is less than the ROPE lower limit, within the ROPE limits, and greater than the ROPE upper limit.

## Exercise 8.3
```{r,warning=FALSE}
##The first line above specifies the beginning of the filenames for saved information
##The second line above specifies the graphics format for saved graphs

fileNameRoot = "Jags-Ydich-XnomSsubj-MbernBeta-Yiru-" 
graphFileType = "eps"

##The MCMC chain is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-Yiru-Mcmc.Rdata
mcmc = genMCMC( data=data , numSavedSteps=10000 , saveName=fileNameRoot )

##The diagnostic graphs are saved in files named Jags-Ydich-XnomSsubj-MbernBeta-Yiru - Diagtheta.eps
parameterNames = varnames(mcmc)
for ( parName in parameterNames ) {
diagMCMC( codaObject=mcmc , parName=parName , saveName=fileNameRoot , saveType=graphFileType )
}

##The summary information is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-Yiru-detail.csv
detail = smryMCMC(mcmc, compVal=0.5, rope=c(0.45,0.55),compValDiff=0.0, ropeDiff = c(-0.05,0.05) ,saveName=fileNameRoot )

##The graph of the posterior distribution is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-Yiru-Post.eps
plotMCMC( mcmc, data=data , compVal=NULL, compValDiff=0.0, saveName=fileNameRoot , saveType=graphFileType)
```  

## Exercise 8.4

### (A)
```{r,warning=FALSE}
source("Jags-Ydich-XnomSsubj-MbernBeta-8.4.R")

fileNameRoot = "Jags-Ydich-XnomSsubj-MbernBeta-Yiru8.4-" 
graphFileType = "eps"

##The MCMC chain is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-Yiru8.4-Mcmc.Rdata
mcmc = genMCMC_8.4( data=data , numSavedSteps=10000 , saveName=fileNameRoot )

##The diagnostic graphs are saved in files named Jags-Ydich-XnomSsubj-MbernBeta-Yiru8.4 - Diagtheta.eps
parameterNames = varnames(mcmc)
for ( parName in parameterNames ) {
diagMCMC( codaObject=mcmc , parName=parName , saveName=fileNameRoot , saveType=graphFileType )
}

##The summary information is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-Yiru8.4-detail.csv
detail = smryMCMC(mcmc, compVal=0.5, rope=c(0.45,0.55),compValDiff=0.0, ropeDiff = c(-0.05,0.05) ,saveName=fileNameRoot )

##The graph of the posterior distribution is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-Yiru8.4-Post.eps
plotMCMC( mcmc, data=data , compVal=NULL, compValDiff=0.0, saveName=fileNameRoot , saveType=graphFileType)
```

### (B)
```{r,warning=FALSE}
source("Jags-Ydich-XnomSsubj-MbernBeta-b.R")

fileNameRoot = "Jags-Ydich-XnomSsubj-MbernBeta-YiruB-" 
graphFileType = "eps"

##The MCMC chain is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-YiruB-Mcmc.Rdata
mcmc = genMCMC_8.4( data=data , numSavedSteps=10000 , saveName=fileNameRoot )

##The diagnostic graphs are saved in files named Jags-Ydich-XnomSsubj-MbernBeta-YiruB - Diagtheta.eps
parameterNames = varnames(mcmc)
for ( parName in parameterNames ) {
diagMCMC( codaObject=mcmc , parName=parName , saveName=fileNameRoot , saveType=graphFileType )
}

##The summary information is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-YiruB-detail.csv
detail = smryMCMC(mcmc, compVal=0.5, rope=c(0.45,0.55),compValDiff=0.0, ropeDiff = c(-0.05,0.05) ,saveName=fileNameRoot )

##The graph of the posterior distribution is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-YiruB-Post.eps
plotMCMC( mcmc, data=data , compVal=NULL, compValDiff=0.0, saveName=fileNameRoot , saveType=graphFileType)
```  
The distributions on theta[1], theta[2] and theta[3] look uniform.

### (C)
```{r,warning=FALSE}
source("Jags-Ydich-XnomSsubj-MbernBeta-c.R")

fileNameRoot = "Jags-Ydich-XnomSsubj-MbernBeta-YiruC-" 
graphFileType = "eps"

##The MCMC chain is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-YiruC-Mcmc.Rdata
mcmc = genMCMC_8.4( data=data , numSavedSteps=10000 , saveName=fileNameRoot )

##The diagnostic graphs are saved in files named Jags-Ydich-XnomSsubj-MbernBeta-YiruC - Diagtheta.eps
parameterNames = varnames(mcmc)
for ( parName in parameterNames ) {
diagMCMC( codaObject=mcmc , parName=parName , saveName=fileNameRoot , saveType=graphFileType )
}

##The summary information is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-YiruC-detail.csv
detail = smryMCMC(mcmc, compVal=0.5, rope=c(0.45,0.55),compValDiff=0.0, ropeDiff = c(-0.05,0.05) ,saveName=fileNameRoot )

##The graph of the posterior distribution is saved in a file named Jags-Ydich-XnomSsubj-MbernBeta-YiruC-Post.eps
plotMCMC( mcmc, data=data , compVal=NULL, compValDiff=0.0, saveName=fileNameRoot , saveType=graphFileType)
```  
The distributions on theta[1], theta[2] and theta[3] look uniform.

