---
title: "Presentation_Reef"
author: "Yiru Fei"
date: "10/20/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
reef <- read.csv("~/Desktop/MSSP/MA615/project/Midterm Project/Reef_Life_Survey_(RLS)#_Global_reef_fish_dataset-fish_surveys.csv")
library(dplyr)
library(ggplot2)
library(kableExtra)
library(knitr)
```

## Introduction

Reef Fish and Water Depth
\newline
This project will define whether the species and amount of reef fish will increase with the depth of sea increase by using the dataset from the Reef Life Survey. 

## Data Description
This data set contains 20 colums, 23655 entries. The following four tables display the first six data of every columns.  

##
<div class="centered">
</div>

```{r, echo = F}
da1 = head(reef[,1:5])
kable(cbind(da1), caption = "First five columns of reef", align = "c", booktab =T, format = "html",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"),position = "center",font_size = 15)
```
##
```{r,echo = F}
da2 = head(reef[,6:10])
kable(cbind(da2), caption = "Second five columns of reef", align = "c", booktab =T, format = "html",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"),position = "center",font_size = 15)
```
##
```{r,echo = F}
da3 = head(reef[,11:15])
kable(cbind(da3), caption = "Third five columns of reef", align = "c", booktab =T, format = "html",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"),position = "center",font_size = 15)
```
##
```{r,echo = F}
da4 = head(reef[,16:20])
kable(cbind(da4), caption = "Last five columns of reef", align = "c", booktab =T, format = "html",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"),position = "center",font_size = 15)
```

## Organize Phylum
<div class="centered">
The most phylum is Chordata in this area.
```{r , echo = F}
reef_phylum  = reef %>% group_by(Depth) %>% count(Phylum)
reef_p = reef %>% count(Phylum)
ggplot(data=reef_p, aes(x = "",y=n, fill = Phylum)) +
  geom_bar(width = 1, stat = "identity")+coord_polar("y", start=0) +labs(title="Phylum")
```  

## Organize Class
<div class="centered">
The most class is Actinopterygii in this area.
</div>
```{r,echo = F}
reef_class = reef %>% group_by(Depth) %>% count(Class)
reef_c = reef %>% group_by(Phylum)%>% count(Class)
kable(reef_c, digits = 2, align = "c", format = "html", booktabs=TRUE, ,caption = "Total Amount of Different Class", col.names = c("Phylum",  "Class" ,"Sum")) %>% kable_styling(latex_options = "hold_position",position = "center",font_size = 15)
```

## Organize Family
<div class="centered">
</div>
  
```{r,echo = F}
reef_family  = reef %>% group_by(Depth) %>% count(Family)
reef_f = reef %>% group_by(Class) %>% count(Family)
reef_f = reef_f[with(reef_f,order(-n)),]
reef_f = head(reef_f,n = 10)
kable(reef_f, digits = 2, align = "c", format = "html", booktabs=TRUE, ,caption = "Top Ten Families", col.names = c("Class","Famliy", "Sum")) %>% kable_styling(latex_options = "hold_position",position = "center",font_size = 15)
```  

## Regression Linear Model
```{r,echo = F}
data <- read.csv("~/Desktop/MSSP/MA615/project/Midterm Project/data1.csv")
data = data[1:89,1:9]

a= lm(Depth~ ., data = data)
par(mfrow=c(2,2))
plot(a)
```

## Transformation
```{r,echo = F}
b= lm(log(Depth)~ ., data = data)
par(mfrow=c(2,2))
plot(b)
```

##
```{r}
b
```
