---
title: "Depth and Reef Fish"
author: "Yiru Fei"
date: "10/20/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
reef <- read.csv("~/Desktop/MSSP/MA615/project/Midterm Project/Reef_Life_Survey_(RLS)#_Global_reef_fish_dataset-fish_surveys.csv")
library(dplyr)
library(ggplot2)
library(kableExtra)
library(knitr)
```

## Introduction

Reef Fish and Water Depth
\newline
This project will define whether the species and amount of reef fish will increase with the depth of sea increase by using the dataset from the Reef Life Survey.  

This project will find the relationship between the species and the amount of reef fish and the depth of the sea. I hypothesis that the species and amount of reef fish will increase as the depth of water increases.

## Data Description
This project used the dataset collected around Australia where belongs to the central Indo-Pacific, and the spatial location is from Lat/Lon $-36.585, 132.429$ to Lat/Lon $-31.18, 139.812$. This dataset contains records of fishes and elasmobranchs collected by a combination of experienced scientists and skilled recreational divers along $50$ m transects on shallow rocky and coral reefs.  

This data set contains 20 colums, 23655 entries. The following four tables display the first six data of every columns. 

##
<div class="centered">
</div>

```{r, echo = F}
da1 = head(reef[,1:5])
kable(cbind(da1), caption = "First five columns of reef", align = "c", booktab =T, format = "html",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"),position = "center",font_size = 15)
```
##
```{r,echo = F}
da2 = head(reef[,6:10])
kable(cbind(da2), caption = "Second five columns of reef", align = "c", booktab =T, format = "html",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"),position = "center",font_size = 15)
```
##
```{r,echo = F}
da3 = head(reef[,11:15])
kable(cbind(da3), caption = "Third five columns of reef", align = "c", booktab =T, format = "html",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"),position = "center",font_size = 15)
```
##
```{r,echo = F}
da4 = head(reef[,16:20])
kable(cbind(da4), caption = "Last five columns of reef", align = "c", booktab =T, format = "html",longtable = F) %>% kable_styling(latex_options = c("HOLD_position"),position = "center",font_size = 15)
```

## Organize Phylum
<div class="centered">
The most phylum is Chordata in this area.
```{r , echo = F}
reef_phylum  = reef %>% group_by(Depth) %>% count(Phylum)
reef_p = reef %>% count(Phylum)
ggplot(data=reef_p, aes(x = "",y=n, fill = Phylum)) +
  geom_bar(width = 1, stat = "identity")+coord_polar("y", start=0) +labs(title="Phylum")
```  

## Organize Class
<div class="centered">
The most class is Actinopterygii in this area.
</div>
```{r,echo = F}
reef_class = reef %>% group_by(Depth) %>% count(Class)
reef_c = reef %>% group_by(Phylum)%>% count(Class)
kable(reef_c, digits = 2, align = "c", format = "html", booktabs=TRUE, ,caption = "Total Amount of Different Class", col.names = c("Phylum",  "Class" ,"Sum")) %>% kable_styling(latex_options = "hold_position",position = "center",font_size = 15)
```

## Organize Family
<div class="centered">
</div>
  
```{r,echo = F}
reef_family  = reef %>% group_by(Depth) %>% count(Family)
reef_f = reef %>% group_by(Class) %>% count(Family)
reef_f = reef_f[with(reef_f,order(-n)),]
reef_f = head(reef_f,n = 10)
kable(reef_f, digits = 2, align = "c", format = "html", booktabs=TRUE, ,caption = "Top Ten Families", col.names = c("Class","Famliy", "Sum")) %>% kable_styling(latex_options = "hold_position",position = "center",font_size = 15)
```  

## Regression Linear Model
```{r,echo = F}
data <- read.csv("~/Desktop/MSSP/MA615/project/Midterm Project/data1.csv")
data = data[1:89,1:9]

a= lm(Depth~ ., data = data)
par(mfrow=c(2,2))
plot(a)
```

## Transformation
```{r,echo = F}
b= lm(log(Depth)~ ., data = data)
par(mfrow=c(2,2))
plot(b)
```

## Linear Regression Model

$log(Depth) = 1.82 - 0.08Chor + 0.13Actino - 0.009Labr -$
$0.04Kyphos + 0.006Mona - 0.03Cheil - 0.02Odac -$
$0.03Taxon$

## Conclusion

In conclusion, the reef fish prefer to live in shallow sea areas, and there is no strong relationship between reef fish and depth since the multiple R-squared is just $0.3$, which means that the points do not fit the regression line well.